Quantizing the Bosonic String on a Loop Quantum Gravity Background

### 2024-06-22 lqg-strings.tex
made some changes in section III

The action eq \eqref{eqn:modified-NG-action} has manifest background Poincaré invariance since just like the regular Nambu-Goto action since all its terms are derivatives of $X^{\mu}$ with all Lorentz indices contracted, thus are Lorentz scalars. At first glance it might seem to not be WS reparameterization invariant because of the second term, however, this is corrected by the fact that $\Delta$ is a scalar \emph{density} of weight 2, thus counteracting the inverse metric determinant. Moreover, if we look at that expression we see that it can be written as:
\begin{equation} \label{eqn:modified-NG-action-v2}
	S'_{NG}[h_{ab}] = -T \int \diff ^2 x \left( \sqrt{- h} - \frac{1}{2} {k \Delta}{\sqrt{- (h^{-1})}} \right),
\end{equation}
where $h^{-1}$ is now the determinant of the \emph{inverse} metric $h^{ab}$. As such, in the expression \eqref{eqn:modified-NG-action} we can see a duality between the first and second terms. If we replace the metric $h_{ab}$ by its inverse $h'_{ab} \equiv h^{ab}$, then the action in terms of $h'_{ab}$ is the same as \eqref{eqn:modified-NG-action}, but with the difference that the factor of $k \Delta / 2$ is mapped to $2 / (k \Delta)$:
\begin{align} \label{eqn:modified-NG-action-v3}
	S'_{NG}[h'_{ab}] & = -T' \int \diff ^2 x' \left( \sqrt{- h'} - \frac{2}{k \Delta} \sqrt{- (h')^{-1}} \right) \nonumber \\
	                 & = -T' \int \diff ^2 x' \left( \sqrt{- h'} - \onehalf {k' \Delta'}\sqrt{- (h')^{-1}} \right),
\end{align}
which has the same form as the original action \eqref{eqn:modified-NG-action}, as long as we make the replacements
\begin{subequations}
	\begin{equation} \label{eqn:T'}
		T \rightarrow T' = \frac{1}{2} k T
	\end{equation}
	\begin{equation} \label{eqn:k-Delta}
		\frac{1}{2} {k \Delta} \rarrow \frac{1}{2} {k' \Delta'} = \frac{2} {k \Delta}
	\end{equation}
	\begin{equation} \label{eqn:d2x}
		\diff ^2 x \rightarrow \diff ^2 x' = \Delta \diff ^2 x
	\end{equation}
\end{subequations}
(or equivalently $\diff x^{a} \rightarrow \diff x' _{a} = \Delta _{a b} \diff x^{b}$). This is a duality between the physics at the Planck scale and the physics at the string scale, \ie between strong coupling ($k' \gg 1$) and weak coupling ($k \ll 1$), small quantum of area ($\Delta$) and large quantum of area ($\Delta'$) and between large string tension $T$ and small string tension $T'$.

@@

Consider a string propagating in a target space whose geometry is described by two bulk metrics. There are two ways that a string with two metrics might arise. One possibility is that there are two sets of embedding fields $X^\mu$ and $Y^\mu$. As a result, the induced metrics on the string worldsheet would also be different for both sets of embedding fields. However, in this case, we would not be talking about \emph{one} string but two \emph{different} strings described by the respective embedding functions. One could then proceed along the lines of \eqref{eqn:hassan-bimetric} and introduce a third term for the worldsheet action which describes a coupling between the two sets of embedding fields, of the form:
\begin{equation}\label{eqn:ws-bimetric-interaction}
	S_{int} = - T_{eff} \int \diff^2 x\, \Phi(g,h) \Psi(\mb{X}, \mb{Y}).
\end{equation}
Here $\Phi(g,h)$ is some function of the \emph{induced} metrics $g_{ab}$ and $h_{ab}$ on the two worldsheets and $\Psi(\mb{X}, \mb{Y})$ is some function of the embedding fields. The resulting action would be of the form:
\begin{equation}\label{eqn:ws-bimetric-action-v1}
	S_{BNG} = -T \int \diff^2 x \, \sqrt{-h} - T' \int \diff^2 x \, \sqrt{-g} + S_{int}.
\end{equation}
Here the subscript $BNG$ stands for ``Bimetric Nambu-Goto''.

@@

In the following, we will derive the equations of motion for the bimetric Polyakov action and along the way we will discover that the scalar density $\Delta(x)$ can be identified as being the determinant of (a tensor proportional to) the antisymmetric tensor $\widetilde{\varepsilon}_{a c}$ living on the worldsheet. The worldsheet antisymmetric tensor can itself be understood as arising from the pullback of a bulk antisymmetric tensor. Now, in string theory, there is already a very important antisymmetric bulk tensor called the Kalb-Ramond field.

Following the usual procedure such as the one found in \cite{Tong2009Lectures}, we find two copies of the string energy-momentum tensor:
\begin{equation} \label{eqn:variation-wrt-g}
	\frac{\delta S_{P B}}{\delta g^{c d}} = 0
\end{equation}
\begin{multline} \label{eqn:g-em-tensor}
	T^{(G)} _{c d} := G_{\mu \nu} \bigg( \partial _{c} X^{\mu} \partial _{d} X^{\nu} - \\
	- \frac{1}{2} g_{c d} g^{a b} \partial _{a} X^{\mu} \partial _{b} X^{\nu} \bigg) = \frac{2}{T} \frac{\delta S_{int}}{\delta g^{c d}} ,
\end{multline}
and completely analogous for the $H_{\mu \nu}$ metric,
\begin{equation} \label{eqn:variation-wrt-h}
	\frac{\delta S_{P B}}{\delta h^{c d}} = 0
\end{equation}
\begin{multline} \label{eqn:h-em-tensor}
	T^{(H)} _{c d} := H_{\mu \nu} \bigg( \partial _{c} X^{\mu} \partial _{d} X^{\mu} - \\
	- \frac{1}{2} h_{c d} h^{a b} \partial _{a} X^{\mu} \partial _{b} X^{\nu} \bigg) = \frac{2}{T'} \frac{\delta S_{int}}{\delta h^{c d}} .
\end{multline}
If we assume no interaction term $S_{int} = 0$, we get the expected expressions for the auxiliary metrics $g_{c d}$ and $h_{c d}$ as in
\begin{equation} \label{eqn:g-expression}
	g_{c d} = 2 f^{(G)} \partial _{c} X^{\mu} \partial _{d} X^{\nu} G_{\mu \nu} ,
\end{equation}
where the function $f^{(G)}$ is given by
\begin{equation} \label{eqn:g's-f-function}
	\frac{1}{f^{(G)}} = g^{a b} \partial _{a} X^{\mu} \partial _{b} X^{\nu} G_{\mu \nu}
\end{equation}
and for $h_{c d}$
\begin{equation} \label{eqn:h-expression}
	h_{c d} = 2 f^{(H)} \partial _{c} X^{\mu} \partial _{d} X^{\nu} H_{\mu \nu} ,
\end{equation}
with $f^{(H)}$ given by
\begin{equation} \label{eqn:h's-f-function}
	\frac{1}{f^{(H)}} = h^{a b} \partial _{a} X^{\mu} \partial _{b} X^{\nu} H_{\mu \nu} .
\end{equation}
If the auxiliary metrics $g_{c d}$ and $h_{c d}$ coincide with the pullback metrics from $G_{\mu \nu}$ and $H_{\mu \nu}$, the action with no interaction reduces to
\begin{equation} \label{eqn:double-NG-action}
	S = - T \int \diff ^2 x \sqrt{- g} - T' \int \diff ^2 x \sqrt{- h} ,
\end{equation}
and from here we note that if $T' = - T k / 2$ and $h = \Delta ^2 / g$, we recover the Nambu-Goto area-corrected action \eqref{eqn:modified-NG-action}
\begin{equation} \label{eqn:area-corrected-NG-from-bimetric}
	S = - T \int \diff ^2 x \left( \sqrt{- g} - \frac{1}{2} \frac{k \Delta}{\sqrt{- g}} \right) .
\end{equation}
The condition $h = \Delta ^2 / g$ can be achieved by
\begin{equation} \label{eqn:h=SgS}
	h_{a b} = \Delta _{a c} \Delta _{b d} g^{c d} .
\end{equation}
This implies that
\begin{multline} \label{eqn:SgS-implication}
	\left( \delta _{a} ^{c'} \delta _{b} ^{d'} H_{\mu \nu} - \frac{1}{2 g} \Delta _{a c} \widetilde{\varepsilon} ^{c c'} \Delta _{b d} \widetilde{\varepsilon} ^{d d'} G_{\mu \nu} \right) \times \\
	\times \del _{c'} X^{\mu} \del _{d'} X^{\nu} = 0 ,
\end{multline}
which can be achieved if
\begin{equation} \label{eqn:S=sigma_y}
	\Delta _{a c} = \sqrt{\Delta} \widetilde{\varepsilon} _{a c}
\end{equation}
and if the background metrics are related by
\begin{equation} \label{eqn:H=1/(2g)G}
	H_{\mu \nu} = \frac{\Delta}{2 g} G_{\mu \nu} .
\end{equation}
This makes sense since $\Delta _{a b}$ is related to area, so it should be a 2-form. Now, since the WS metric $g_{a b}$ is taken as the pullback of the bulk metric $g_{a b} = \del _{a} X^{\mu} \del _{b} X^{\nu} G_{\mu \nu}$, which is part of the closed string spin-2 spectrum, then it would make sense for the WS area 2-form $\Delta _{a b}$ to be (at least related to) the pullback of the Kalb-Ramond field, which is the other half of the closed string rank 2 tensor field excitations,
\begin{equation} \label{eqn:Delta-B-relation}
	\Delta _{a b} \sim \del _{a} X^{\mu} \del _{b} X^{\nu} B_{\mu \nu} .
\end{equation}

@@

The equations of motion for the world-sheet metric are completely unchanged by this correction,
\begin{equation} \label{eqn:vary-wrt-g}
	\frac{\delta S_{I A P}}{\delta g^{c d}} = 0
\end{equation}
which leads to the stress-energy tensor for the worldsheet:
\begin{multline} \label{eqn:inverse-area-em-tensor}
	T_{c d} := \partial _{c} X^{\mu} \partial _{d} X^{\nu} G_{\mu \nu} - \\ - \frac{1}{2} g_{c d} g^{a b} \partial _{a} X^{\mu} \partial _{b} X^{\nu} G_{\mu \nu} = 0 ,
\end{multline}
from which we can extract the metric $g_{c d}$ as
\begin{equation} \label{eqn:ws-metric}
	g_{c d} = 2 f \partial _{c} X^{\mu} \partial _{d} X^{\nu} G_{\mu \nu} ,
\end{equation}
where the function $f$ is given by
\begin{equation} \label{eqn:inverse-area-f}
	\frac{1}{f} = g^{a b} \partial _{a} X^{\mu} \partial _{b} X^{\nu} G_{\mu \nu}
\end{equation}
just like the regular ST literature.

>>>>

The action eq \eqref{eqn:modified-NG-action} has manifest background Poincaré invariance since just like the regular Nambu-Goto action since all its terms are derivatives of $X^{\mu}$ with all Lorentz indices contracted, thus are Lorentz scalars. At first glance it might seem to not be WS reparameterization invariant because of the second term, however, this is corrected by the fact that $\Delta$ is a scalar \emph{density} of weight 2, thus counteracting the inverse metric determinant. Moreover, if we look at that expression we see that it can be written as:
\begin{equation} \label{eqn:modified-NG-action-v2}
	S'_{NG}[h_{ab}] = -T \int \diff ^2 x \left( \sqrt{- h} - \frac{1}{2} {k \Delta}{\sqrt{- (h^{-1})}} \right),
\end{equation}
where $h^{-1}$ is now the determinant of the \emph{inverse} metric $h^{ab}$. As such, in the expression \eqref{eqn:modified-NG-action} we can see a duality between the first and second terms. If we replace the metric $h_{ab}$ by its inverse $h'_{ab} \equiv h^{ab}$, then the action in terms of $h'_{ab}$ is the same as \eqref{eqn:modified-NG-action}, but with the difference that the factor of $k \Delta / 2$ is mapped to $2 / (k \Delta)$:
\begin{align} \label{eqn:modified-NG-action-v3}
	S'_{NG}[h'_{ab}] & = -T' \int \diff ^2 x' \left( \sqrt{- h'} - \frac{2}{k \Delta} \sqrt{- (h')^{-1}} \right) \nonumber \\
	                 & = -T' \int \diff ^2 x' \left( \sqrt{- h'} - \onehalf {k' \Delta'}\sqrt{- (h')^{-1}} \right),
\end{align}
which has the same form as the original action \eqref{eqn:modified-NG-action}, as long as we make the replacements
\begin{subequations}
	\begin{equation} \label{eqn:T'}
		T \rightarrow T' = - \frac{1}{2} k T
	\end{equation}
	\begin{equation} \label{eqn:k-Delta}
		\frac{1}{2} {k \Delta} \rarrow \frac{1}{2} {k' \Delta'} = \frac{2} {k \Delta}
	\end{equation}
	\begin{equation} \label{eqn:d2x}
		\diff ^2 x \rightarrow \diff ^2 x' = \Delta \diff ^2 x
	\end{equation}
\end{subequations}
(or equivalently $\diff x^{a} \rightarrow \diff x' _{a} = \Delta _{a b} \diff x^{b}$). This is a duality between the physics at the Planck scale and the physics at the string scale, \ie between strong coupling ($k' \gg 1$) and weak coupling ($k \ll 1$), small quantum of area ($\Delta$) and large quantum of area ($\Delta'$) and between large string tension $T$ and small string tension $T'$.

@@

There are two ways that a string with two metrics might arise. One possibility is that there are two sets of embedding fields $X^\mu$ and $Y^\mu$. As a result, the induced metrics on the string worldsheet would also be different for both sets of embedding fields. However, in this case, we would not be talking about \emph{one} string but two \emph{different} strings described by the respective embedding functions. One could then proceed along the lines of \eqref{eqn:hassan-bimetric} and introduce a third term for the worldsheet action which describes a coupling between the two sets of embedding fields, of the form:
\begin{equation}\label{eqn:ws-bimetric-interaction}
	S_{int} = - T_{eff} \int \diff^2 x\, \Phi(g,h) \Psi(\mb{X}, \mb{Y}).
\end{equation}
Here $\Phi(g,h)$ is some function of the \emph{induced} metrics $g_{ab}$ and $h_{ab}$ on the two worldsheets and $\Psi(\mb{X}, \mb{Y})$ is some function of the embedding fields. The resulting action would be of the form:
\begin{equation}\label{eqn:ws-bimetric-action-v1}
	S_{BNG} = -T \int \diff^2 x \, \sqrt{-h} - T' \int \diff^2 x \, \sqrt{-g} + S_{int}.
\end{equation}
Here the subscript $BNG$ stands for ``Bimetric Nambu-Goto''.

@@

In the following, we will derive the equations of motion for the bimetric Polyakov action and along the way we will discover that the scalar density $\Delta(x)$ can be identified as being the determinant of (a tensor proportional to) the antisymmetric symbol $\varepsilon _{a c}$ living on the worldsheet. The worldsheet antisymmetric symbol can itself be understood as arising from the pullback of a bulk antisymmetric tensor. Now, in string theory, there is already a very important antisymmetric background tensor: the Kalb-Ramond field.

Following the usual procedure such as the one found in \cite{Tong2009Lectures}, we find two copies of the string energy-momentum tensor:
\begin{equation} \label{eqn:variation-wrt-g}
	\frac{\delta S_{P B}}{\delta g^{c d}} = 0
\begin{multline} \label{eqn:g-em-tensor}
	T^{(G)} _{c d} := G_{\mu \nu} \bigg( \partial _{c} X^{\mu} \partial _{d} X^{\nu} - \\
	- \frac{1}{2} g_{c d} g^{a b} \partial _{a} X^{\mu} \partial _{b} X^{\nu} \bigg) = \frac{2}{T} \frac{\delta S_{int}}{\delta g^{c d}} ,
and completely analogous for the $H_{\mu \nu}$ metric,
\begin{equation} \label{eqn:variation-wrt-h}
	\frac{\delta S_{P B}}{\delta h^{c d}} = 0
\begin{multline} \label{eqn:h-em-tensor}
	T^{(H)} _{c d} := H_{\mu \nu} \bigg( \partial _{c} X^{\mu} \partial _{d} X^{\mu} - \\
	- \frac{1}{2} h_{c d} h^{a b} \partial _{a} X^{\mu} \partial _{b} X^{\nu} \bigg) = \frac{2}{T'} \frac{\delta S_{int}}{\delta h^{c d}} .
If we assume no interaction term $S_{int} = 0$, we get the expected expressions for the auxiliary metrics $g_{c d}$ and $h_{c d}$ as in
\begin{equation} \label{eqn:g-expression}
	g_{c d} = 2 f^{(G)} \partial _{c} X^{\mu} \partial _{d} X^{\nu} G_{\mu \nu} ,
where the function $f^{(G)}$ is given by
\begin{equation} \label{eqn:g's-f-function}
	\frac{1}{f^{(G)}} = g^{a b} \partial _{a} X^{\mu} \partial _{b} X^{\nu} G_{\mu \nu}
and for $h_{c d}$
\begin{equation} \label{eqn:h-expression}
	h_{c d} = 2 f^{(H)} \partial _{c} X^{\mu} \partial _{d} X^{\nu} H_{\mu \nu} ,
with $f^{(H)}$ given by
\begin{equation} \label{eqn:h's-f-function}
	\frac{1}{f^{(H)}} = h^{a b} \partial _{a} X^{\mu} \partial _{b} X^{\nu} H_{\mu \nu} .
If the auxiliary metrics $g_{c d}$ and $h_{c d}$ coincide with the pullback metrics from $G_{\mu \nu}$ and $H_{\mu \nu}$, the action with no interaction reduces to
\begin{equation} \label{eqn:double-NG-action}
	S = - T \int \diff ^2 x \sqrt{- g} - T' \int \diff ^2 x \sqrt{- h} ,
and from here we note that if $T' = - T k / 2$ and $h = \Delta ^2 / g$, we recover the Nambu-Goto area-corrected action \eqref{eqn:modified-NG-action}
\begin{equation} \label{eqn:area-corrected-NG-from-bimetric}
	S = - T \int \diff ^2 x \left( \sqrt{- g} - \frac{1}{2} \frac{k \Delta}{\sqrt{- g}} \right) .
The condition $h = \Delta ^2 / g$ can be achieved by
\begin{equation} \label{eqn:h=SgS}
	h_{a b} = \Delta _{a c} \Delta _{b d} g^{c d} .
By writing the inverse metric $g^{a b}$ in terms of Levi-Civita symbols $g^{a b} = \varepsilon ^{a c} \varepsilon ^{b d} g_{c d} / 2 g$, this implies that
\begin{multline} \label{eqn:SgS-implication}
	\left( \delta _{a} ^{c'} \delta _{b} ^{d'} H_{\mu \nu} - \frac{1}{2 g} \Delta _{a c} \varepsilon ^{c c'} \Delta _{b d} \varepsilon ^{d d'} G_{\mu \nu} \right) \times \\
	\times \del _{c'} X^{\mu} \del _{d'} X^{\nu} = 0 ,
which can be achieved if the expression in parenthesis vanishes. For this end, the most natural possibility is
\begin{equation} \label{eqn:S=sigma_y}
	\Delta _{a c} = \sqrt{\Delta} \varepsilon _{a c}
since we must have that $\det (\Delta _{a b}) = \Delta$ and $\Delta _{a c} \varepsilon ^{c c'} \sim \delta ^{c'} _{a}$, and also if the background metrics are related by
\begin{equation} \label{eqn:H=1/(2g)G}
	H_{\mu \nu} = \frac{\Delta}{2 g} G_{\mu \nu} .
This makes sense since $\Delta _{a b}$ is related to area, so it should be a 2-form. Now, since the WS metric $g_{a b}$ is taken as the pullback of the bulk metric $g_{a b} = \del _{a} X^{\mu} \del _{b} X^{\nu} G_{\mu \nu}$, which is part of the closed string spin-2 spectrum, then it would make sense for the WS area 2-form $\Delta _{a b}$ to be (at least related to) the pullback of the Kalb-Ramond field, which is the other half of the closed string rank 2 tensor field excitations,
\begin{equation} \label{eqn:Delta-B-relation}
	\Delta _{a b} \sim \del _{a} X^{\mu} \del _{b} X^{\nu} B_{\mu \nu} .

@@

The equations of motion for the world-sheet metric are completely unchanged by this correction,
\begin{equation} \label{eqn:vary-wrt-g}
	\frac{\delta S_{I A P}}{\delta g^{c d}} = 0
\end{equation}
which leads to the stress-energy tensor for the worldsheet:
\begin{multline} \label{eqn:inverse-area-em-tensor}
	T_{c d} := \partial _{c} X^{\mu} \partial _{d} X^{\nu} G_{\mu \nu} - \\ - \frac{1}{2} g_{c d} g^{a b} \partial _{a} X^{\mu} \partial _{b} X^{\nu} G_{\mu \nu} = 0 ,
\end{multline}
from which we can extract the metric $g_{c d}$ as
\begin{equation} \label{eqn:ws-metric}
	g_{c d} = 2 f \partial _{c} X^{\mu} \partial _{d} X^{\nu} G_{\mu \nu} ,
\end{equation}
where the function $f$ is given by
\begin{equation} \label{eqn:inverse-area-f}
	\frac{1}{f} = g^{a b} \partial _{a} X^{\mu} \partial _{b} X^{\nu} G_{\mu \nu}
\end{equation}
just like the regular ST literature, thus the constraints are unchanged by our correction.
